---
title: "In-Class Exercise 3: Kernel Density Estimation"
author:
  - name: Matthew Ho
    url: https://www.linkedin.com/in/matthewhoyiwen/
date: 01-22-2024
description: |
  In-class exercise 3 
categories:
  - In-class Exercise
format:
  html:
    toc: true
execute: 
  eval: true
  echo: true
  warning: false
---

## 0.0 Class Notes

#### Spatial Point Patterns

-   Points as events, relates to activities

-   Issues with points: Highly affected by the size of the 'dots'. KDE is more consistent

-   The points MUST be a population. Cannot be a sample or we will have sample bias.

#### KDE Setup

-   Choose bandwidth
-   Choose a kernel

#### NKDE: Network-Constrained Kernel Density Estimation

-   Constrained by network

#### TNKDE: Temporal Network Kernel Density Estimate

-   Accounts for both time and location

## 1.0 Setup

### 1.1 Dependencies

```{r}
pacman::p_load(tidyverse, sf, tmap, maptools, spatstat, raster)
```

### 1.2 Importing and Changing CRS

```{r}
childcare_sf <- st_read("data/geospatial/ChildCareServices.geojson") %>% 
  st_transform(crs=3414)

mpsz_sf <- st_read(dsn = "data/geospatial/", layer="MP14_SUBZONE_WEB_PL") %>%
  st_transform(crs=3414)

st_crs(childcare_sf)
st_crs(mpsz_sf)
```

### 1.3 Deriving Coastal Outline

```{r}
sg_sf <- st_union(mpsz_sf)
```

#### Important: 1 Liner conversion

```{r}
childcare_ppp <- as.ppp(childcare_sf)
summary(childcare_ppp)
```

### 1.4 Removing Duplicates

```{r}
any(duplicated(childcare_ppp))
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)

any(duplicated(childcare_ppp_jit))
```

#### Another one-liner

```{r}
sg_owin <- as.owin(sg_sf, "owin")
sg_owin
```

## 2.0 Simple Visualisations

```{r}
plot(mpsz_sf)
```

```{r}
tmap_mode('view')
tm_shape(childcare_sf) +
  tm_dots(alpha=0.4, 
          size=0.05)
tmap_mode('plot')
```

### 2.1 Plot the ppp object within the owin

```{r}
plot(childcare_ppp[sg_owin])
```

### 2.2 Using the filter method

```{r}
pg = mpsz_sf %>% filter(PLN_AREA_N == "PUNGGOL")
tm = mpsz_sf %>% filter(PLN_AREA_N == "TAMPINES")
ck = mpsz_sf %>% filter(PLN_AREA_N == "CHOA CHU KANG")
jw = mpsz_sf %>% filter(PLN_AREA_N == "JURONG WEST")
```

#### Extras

```{r fig.width=16, fig.height=10}
par(mfrow=c(2,2))
plot(pg["SUBZONE_N"])
plot(tm["SUBZONE_N"])
plot(ck["SUBZONE_N"])
plot(jw["SUBZONE_N"])
```
