---
title: "In-Class Exercise 3.1: Kernel Density Estimation"
author:
  - name: Matthew Ho
    url: https://www.linkedin.com/in/matthewhoyiwen/
date: 01-22-2024
description: |
  In-class exercise 3.1
categories:
  - In-class Exercise
format:
  html:
    toc: true
execute: 
  eval: true
  echo: true
  warning: false
---

## 0.0 Class Notes

#### Spatial Point Patterns

-   Points as events, relates to activities

-   Issues with points: Highly affected by the size of the 'dots'. KDE is more consistent

-   The points MUST be a population. Cannot be a sample or we will have sample bias.

#### KDE Setup

-   Choose bandwidth
-   Choose a kernel

#### NKDE: Network-Constrained Kernel Density Estimation

-   Constrained by network

#### TNKDE: Temporal Network Kernel Density Estimate

-   Accounts for both time and location

## 1.0 Setup

### 1.1 Dependencies

```{r}
pacman::p_load(sf, spNetwork, tmap, classInt, viridis, tidyverse)
```

### 1.2 Importing and Changing CRS

```{r}
network <- st_read(dsn="data/geospatial", layer="Punggol_St")
childcare <- st_read(dsn="data/geospatial", layer="Punggol_CC")
```

```{r}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots() +
  tm_shape(network) +
  tm_lines()
tmap_mode('plot')

```