---
title: "IS415 Hands-On Exercise 1"
---

Use `pacman::p_load` to install and load sf and tidyverse

```{r}
pacman::p_load(sf, tidyverse)
```

---

`st_read` to open and import the **shapefile**.

-   Path to **shapefile**

-   **Shapefile** without the file extension

```{r}
mpsz = st_read(dsn = "data/geospatial", 
               layer = "MP14_SUBZONE_WEB_PL")
cyclingpath = st_read(dsn = "data/geospatial", layer = "CyclingPathGazette")
preschool = st_read("data/geospatial/PreSchoolsLocation.kml")
```

---

`st_geometry` prints basic info on the feature class

```{r}
st_geometry(mpsz)
```

---

`glimpse` shows the associated attribute info

```{r}
glimpse(mpsz)
```

---

`head` shows complete info on feature object

```{r}
head(mpsz, n=5)  
```

---

`plot` a multi-plot of all attributes (default for sf object)

```{r}
plot(mpsz, max.plot = 15)
```

---

Plot just the geometry

```{r}
plot(mpsz["PLN_AREA_N"])
```

---

Show the **EPSG** code of simple feature data frame (See last, CS, section). It is 9001.

```{r}
st_crs(mpsz)
```

---

Change it to **SVY21** (3414)

```{r}
mpsz3414 <- st_transform(mpsz, 3414)
st_crs(mpsz3414)
preschool3414 <- st_transform(preschool, crs = 3414)
st_crs(preschool3414)
```

---

**Aspatial** data: not geospatial data.

```{r}
listings <- read_csv("data/aspatial/listings.csv")
list(listings) 
```

---

Convert the longitude and latitude columns from **WGS84** to **SVY21**.

```{r}
listings_sf <- st_as_sf(listings, 
                        coords = c("longitude", "latitude"),
                        crs=4326) %>%
  st_transform(crs = 3414)

glimpse(listings_sf)
```

---

**Buffering**: Add space along the side.

-   5 Meter buffers

```{r}
buffer_cycling <- st_buffer(cyclingpath, 
                            dist=5, nQuadSegs = 30)
```

---

Find the area of the buffers, i.e. see how much space is needed.

```{r}
buffer_cycling$AREA <- st_area(buffer_cycling)
sum(buffer_cycling$AREA)
```

---

**Point in Polygon Count**: How many points in a polygon. E.g. count preschools in each zone.

-   Use `st_intersects` for this

-   NOT `st_intersection`

```{r}
mpsz3414$`PreSch Count`<- lengths(st_intersects(mpsz3414, preschool3414))
```

---

Print the summary stats

```{r}
summary(mpsz3414$`PreSch Count`)
```

---

`top_n` to show the rows with the highest values in a column

```{r}
top_n(mpsz3414, 3, `PreSch Count`)
```

---

Calculate area of each subzone and **mutate** to compute density

```{r}
mpsz3414$Area <- mpsz3414 %>%
  st_area()
mpsz3414 <- mpsz3414 %>%
  mutate(`PreSch Density` = `PreSch Count`/Area * 1000000)
```

---

EDA Histogram

```{r}
hist(mpsz3414$`PreSch Density`)
```

---

`ggplot2` makes cooler graphics

```{r}
ggplot(data=mpsz3414, 
       aes(x= as.numeric(`PreSch Density`)))+
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  labs(title = "Are pre-school even distributed in Singapore?",
       subtitle= "There are many planning sub-zones with a single pre-school, on the other hand, \nthere are two planning sub-zones with at least 20 pre-schools",
       x = "Pre-school density (per km sq)",
       y = "Frequency")
```

---

Scatterplot of pre-school count vs pre-school density

```{r}
ggplot(data=mpsz3414, 
       aes(y = `PreSch Count`, 
           x= as.numeric(`PreSch Density`))) +
  geom_point(color="black", 
             fill="light blue") +
  xlim(0, 40) +
  ylim(0, 40) +
  labs(title = "",
       x = "Pre-school density (per km sq)",
       y = "Pre-school count")

```
