---
title: "Take-Home Exercise 3"
author:
  - name: Matthew Ho
    url: https://www.linkedin.com/in/matthewhoyiwen/
date: 03-10-2024
description: |
  Take-Home Exercise 3
categories:
  - Take-Home Exercise
format:
  html:
    toc: true
execute: 
  eval: true
  echo: true
  warning: false
---

## Overview and Objectives

In this takehome exercise, I will be prototyping my accessibility modelling module for the Geospatial Analytics Project. 

This module aims to 
 -    Model the accessibility to various amenties across Singapore
 -    Show the distributions of accessibility to amenities across regions
 
 -    Choose between Hexagonal and Square Grids
 -    Change the desired grid size
 -    Choose different distance decay functions
 -    Choose between Hansen’s, KD2SFCA and SAM accessibility modelling methods

## Data Wrangling

```{r}
pacman::p_load(tidyverse, tmap, sf, dplyr, smoothr, tmap, SpatialAcc, hash, ggplotify, cowplot)
```

```{r}
mpsz <- st_read("data/geospatial/MasterPlan2019SubzoneBoundaryNoSeaKML.kml") %>% 
  mutate(Match=str_match_all(Description,"<td>(.*?)</td>")) %>% 
  mutate(Match=map(Match, ~ .[,2])) %>% 
  mutate(Match=map(Match,setNames,c("SUBZONE_NO","SUBZONE_N","SUBZONE_C", "CA_IND", "PLN_AREA_N", "PLN_AREA_C", "REGION_N", "REGION_C", "INC_CRC", "FMEL_UPD_D"))) %>% 
  unnest_wider(Match) %>%
  st_as_sf() %>% dplyr::select('Name', 'geometry', 'SUBZONE_NO', 'SUBZONE_N', 'PLN_AREA_N', 'PLN_AREA_C', 'REGION_N', 'REGION_C') %>% 
  st_make_valid() %>%
  st_zm() %>%
  st_transform(crs = 3414) %>%
  write_rds('data/rds/mpsz.rds')
```
::: panel-tabset
## Planning Area Level
```{r}
pop2020_pa <- read_rds('data/rds/pop2020.rds') %>%
  mutate(Total = as.numeric(Total)) %>%
  group_by(Planning_Area) %>%
  summarise(Total = sum(Total), .groups = "drop")

mpsz_pa <- mpsz %>%
  group_by(PLN_AREA_N) %>%
  summarise(geometry = st_union(geometry), .groups = "drop") %>% 
  merge(pop2020_pa, by.x="PLN_AREA_N", by.y="Planning_Area") %>% 
  mutate(Total = as.numeric(Total),
           area = st_area(geometry), 
           pop_dens = Total / area)

mpsz_pa$area <- st_area(mpsz_pa$geometry)
mpsz_pa$pop_dens <- mpsz_pa$Total / mpsz_pa$area

write_rds(mpsz_pa, 'data/rds/mpsz_pa.rds')
```

## Subzone Level

```{r}
pop2020_sz <- read_rds('data/rds/pop2020.rds') %>%
  mutate(Total = as.numeric(Total)) %>%
  group_by(Subzone) %>%
  summarise(Total = sum(Total), .groups = "drop")

mpsz_sz <- mpsz %>%
  group_by(SUBZONE_N) %>%
  summarise(geometry = st_union(geometry), .groups = "drop") %>% 
  merge(pop2020_sz, by.x="SUBZONE_N", by.y="Subzone") %>% 
  mutate(Total = as.numeric(Total),
           area = st_area(geometry), 
           pop_dens = Total / area)

mpsz_sz$area <- st_area(mpsz_sz$geometry)
mpsz_sz$pop_dens <- mpsz_sz$Total / mpsz_sz$area

write_rds(mpsz_sz, 'data/rds/mpsz_sz.rds')
```

:::

#### Getting Mainland Singapore
```{r}
mainland_sg <- st_union(mpsz_pa) %>%
    st_cast("POLYGON")

mainland_sg <- mainland_sg[c(15)] %>% 
  fill_holes(units::set_units(1, "km^2")) %>%
  st_as_sf() %>%
  write_rds('data/rds/mainland_sg.rds')

plot(mainland_sg)
```

## Grid Generation Process

#### Function to add "demand" 
```{r}
add_weights <- function(grids, pop) {
  joined <- st_join(grids, pop, join = st_intersects)  %>%
    mutate(intersect_area = st_area(x),
           demand = intersect_area * pop_dens) %>%
    st_drop_geometry()
  
  grids <- grids %>%
    left_join(joined %>%
                group_by(ID) %>%
                summarise(total_demand = sum(demand)),
              by = "ID") %>%
    mutate(demand = ifelse(is.na(total_demand), 0, total_demand)) %>%
    select(-total_demand)
}
```

#### Make a grid over mainland Singapore
```{r}
sg_2_grid <- function(sg, g_size, is_square, name, pop) {
  grids <- st_make_grid(sg, cellsize = g_size, square = is_square) %>%
  st_intersection(mainland_sg) %>%
  st_as_sf() %>%
    mutate(demand = 0, ID = row_number())
  
  grids_w_demand <- add_weights(grids, pop) %>%
  rename(geometry = x)
  
  write_rds(grids_w_demand, paste(name, '.rds', sep = ""))
  
  return(grids_w_demand)
}
```

#### Generate the grids and distance matrices
```{r}
grids_and_dist_matrices <- function(sg, pop, g_size, is_square, subz) {
  name <- paste('data/rds/grid_', g_size, ifelse(is_square, "_square", "_hexagon"), ifelse(subz, "_sz", "_pa"), sep = "")
  grid <- sg_2_grid(sg, g_size, is_square, name, pop)
  return(grid)
}
```

#### Writing the grids with population as demand
```{r}
#| code-fold: true
#| code-summary: "Show the code"

# for(di in c(250, 500, 750, 1000)) {
#   grids_and_dist_matrices(mainland_sg, mpsz_sz, di, FALSE, TRUE)
#   grids_and_dist_matrices(mainland_sg, mpsz_sz, di, TRUE, TRUE)
#   grids_and_dist_matrices(mainland_sg, mpsz_pa, di, FALSE, FALSE)
#   grids_and_dist_matrices(mainland_sg, mpsz_pa, di, TRUE, FALSE)
# }
```

### Functional Demo
```{r}
demo <- grids_and_dist_matrices(mainland_sg, mpsz_sz, 500, FALSE, TRUE)

tm_shape(demo) + 
  tm_fill(col='demand',
            n = 7,
            style = "quantile",
            border.col = "black",
            border.lwd = 1,
            na.rm = TRUE)
```
## Step by Step Demo

#### Generating the grid
```{r}
grids <- st_make_grid(mainland_sg, cellsize = 500, square = FALSE)
plot(grids)
```

```{r}
grids <- grids %>%
st_intersection(mainland_sg) %>%
st_as_sf() %>%
  mutate(demand = 0, ID = row_number())

tm_shape(grids) + 
  tm_polygons(col='demand',
            n = 7,
            style = "quantile",
            border.col = "black",
            border.lwd = 1,
            na.rm = TRUE)
```
```{r}
tm_shape(mpsz_sz) + 
  tm_polygons(col='Total',
            n = 7,
            style = "quantile",
            border.col = "black",
            border.lwd = 1,
            na.rm = TRUE)
```

```{r}
grids_on_sz <- st_join(grids, mpsz_sz, join = st_intersects)  %>%
  mutate(intersect_area = st_area(x),
         demand = intersect_area * pop_dens) %>%
  st_drop_geometry()

grids <- grids %>%
  left_join(grids_on_sz %>%
              group_by(ID) %>%
              summarise(total_demand = sum(demand)),
            by = "ID") %>%
  mutate(demand = ifelse(is.na(total_demand), 0, total_demand)) %>%
  select(-total_demand) %>%
rename(geometry = x)
```
```{r}
tm_shape(grids) + 
  tm_polygons(col='demand',
            n = 7,
            style = "quantile",
            border.col = "black",
            na.rm = TRUE)
```
### Handling Points

::: panel-tabset

## Supermarkets
```{r}
supermarkets <- st_read("data/geospatial/SupermarketsKML.kml") %>% 
  mutate(Match=str_match_all(Description,"<td>(.*?)</td>")) %>% 
  mutate(Match=map(Match, ~ .[,2])) %>% 
  mutate(Match=map(Match,setNames,c("LIC_NAME", "BLK_HOUSE", "STR_NAME", "UNIT_NO", "POSTCODE", "LIC_NO", "INC_CRC", "FMEL_UPD_D"))) %>% 
  unnest_wider(Match) %>%
  st_as_sf() %>% 
  dplyr::select('LIC_NAME', 'geometry') %>% 
  st_make_valid() %>%
  st_zm() %>%
  st_transform(crs = 3414)
```

```{r}
any(is.na(supermarkets))
```

```{r}
supermarkets$SUBTYPE <- ifelse(grepl("FAIRPRICE", supermarkets$LIC_NAME, ignore.case = TRUE), "FAIRPRICE",
                     ifelse(grepl("COLD STORAGE", supermarkets$LIC_NAME, ignore.case = TRUE), "COLD STORAGE",
                            ifelse(grepl("SHENG SIONG", supermarkets$LIC_NAME, ignore.case = TRUE), "SHENG SIONG", "OTHER")))

write_rds(supermarkets, 'data/rds/supermarkets.rds')
```

```{r}
tm_shape(mainland_sg) + 
  tm_polygons() + 
  tm_shape(supermarkets) + 
  tm_dots(col = "SUBTYPE", size = 0.05)
```

## Markets and Food Centres

```{r}
marketandfc <- st_read("data/geospatial/NEAMarketandFoodCentreKML.kml") %>% 
  mutate(Match=str_match_all(Description,"<td>(.*?)</td>")) %>% 
  mutate(Match=map(Match, ~ .[,2])) %>% 
  mutate(Match=map(Match,setNames,c("TOTAL_STALLS", "MP_STALLS", "CF_STALLS", "POSTAL_CODE", "OWNER", "TYPE", "LOCATION_CENTRE", "NAME_OF_CENTRE", "INC_CRC", "FMEL_UP_D"))) %>% 
  unnest_wider(Match) %>%
  st_as_sf() %>% dplyr::select("NAME_OF_CENTRE", "TYPE", "TOTAL_STALLS", "geometry") %>% 
  st_make_valid() %>%
  st_zm() %>%
  st_transform(crs = 3414)
```

```{r}
any(is.na(marketandfc))
```

```{r}
marketandfc$SUBTYPE <- ifelse(marketandfc$TYPE == "HC", "HAWKER_CENTRE",
                  ifelse(marketandfc$TYPE == "MHC", "MARKET_AND_HAWKER",
                         ifelse(marketandfc$TYPE == "MK", "MARKET", NA)))
marketandfc <- subset(marketandfc, select = -TYPE)

marketandfc$TOTAL_STALLS <- as.numeric(marketandfc$TOTAL_STALLS)

write_rds(marketandfc, 'data/rds/markets_and_food_centres.rds')
```

```{r}
tm_shape(mainland_sg) + 
  tm_polygons() + 
tm_shape(marketandfc) + 
  tm_dots(col = "SUBTYPE", group = "SUBTYPE", size = "TOTAL_STALLS")
```


## MRT

```{r}
mrt <- read.csv('data/aspatial/mrt.csv') %>%
  st_as_sf(coords = c("lng", "lat"),
                      crs = 4326) %>%
  st_transform(crs = 3414) %>%
  dplyr::select('name', 'geometry')

write_rds(mrt, 'data/rds/mrt.rds')
```

```{r}
any(is.na(mrt))
```
```{r}
tm_shape(mainland_sg) + 
  tm_polygons() + 
tm_shape(mrt) + 
  tm_dots(col = "red")
```
## Other POIs

```{r}
poi <- read.csv('data/aspatial/poi.csv') %>%
  st_as_sf(coords = c("lng", "lat"),
                      crs = 4326) %>%
  st_transform(crs = 3414)
```

```{r}
poi2type <- function(pois, typ) {
  poi <- pois[pois[[typ]] == 'True', ] %>%
    dplyr::select('name', 'geometry')
  write_rds(poi, paste('data/rds/poi_', typ, '.rds', sep=""))
}
```

```{r}
poi_types <- c("restaurant", "hospital", "lodging", "bank", "cafe", "convenience_store", "clothing_store", "atm", "school", "beauty_salon", "place_of_worship", "tourist_attraction", "doctor", "dentist", "gym", "night_club", "library")

for (poi_t in poi_types) {
  poi2type(poi, poi_t)
}
```

:::

## Accessibility Modelling

```{r}
grid <- read_rds('data/rds/grid_500_hexagon_sz.rds')
points <- read_rds('data/rds/supermarkets.rds') %>%
  mutate(capacity = 500)
```

::: panel-tabset

## Hansen Method
```{r}
#| code-fold: true
#| code-summary: "Show the code"

exponent <- 2
method <- "Hansen"

centroid.coords <- st_coordinates(st_centroid(grid))
points.coords <- st_coordinates(points)

dm <- exp(distance(centroid.coords, points.coords, type = "euclidean") / 1000 * exponent)

acc <- data.frame(ac(grid$demand,
                          points$capacity,
                          dm, 
                          d0 = 250,
                          power = 2, 
                          family = method))

colnames(acc) <- "acc"
hexagon <- bind_cols(grid, as_tibble(acc))
hexagon$acc[is.infinite(hexagon$acc)] <- NA

mapex <- st_bbox(grid)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"

  tm_shape(grid) + 
  tm_polygons() +
  tm_shape(hexagon,
           bbox = mapex) + 
    tm_fill(col = "acc",
            n = 5,
            style = "quantile",
            border.col = "black",
            border.lwd = 1,
            na.rm = TRUE) +
  tm_shape(points) +
    tm_symbols(size = 0.1) +
    tm_layout(main.title = paste("Accessibility to supermarkets: Hansen method", sep=""),
              main.title.position = "center",
              main.title.size = 2,
              legend.outside = FALSE,
              legend.height = 0.45, 
              legend.width = 3.0,
              legend.format = list(digits = 6),
              legend.position = c("right", "top"),
              frame = TRUE) +
    tm_compass(type="8star", size = 2) +
    tm_scale_bar(width = 0.15) +
    tm_grid(lwd = 0.1, alpha = 0.5)
```

## KD2SFCA Method
```{r}
#| code-fold: true
#| code-summary: "Show the code"

exponent <- 2
method <- "KD2SFCA"

centroid.coords <- st_coordinates(st_centroid(grid))
points.coords <- st_coordinates(points)

dm <- exp(distance(centroid.coords, points.coords, type = "euclidean") / 1000 * exponent)

acc <- data.frame(ac(grid$demand,
                          points$capacity,
                          dm, 
                          d0 = 250,
                          power = 2, 
                          family = method))

colnames(acc) <- "acc"
hexagon <- bind_cols(grid, as_tibble(acc))
hexagon$acc[is.infinite(hexagon$acc)] <- NA

mapex <- st_bbox(grid)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"

  tm_shape(grid) + 
  tm_polygons() +
  tm_shape(hexagon,
           bbox = mapex) + 
    tm_fill(col = "acc",
            n = 5,
            style = "quantile",
            border.col = "black",
            border.lwd = 1,
            na.rm = TRUE) +
  tm_shape(points) +
    tm_symbols(size = 0.1) +
    tm_layout(main.title = paste("Accessibility to supermarkets: KD2FCSA method", sep=""),
              main.title.position = "center",
              main.title.size = 2,
              legend.outside = FALSE,
              legend.height = 0.45, 
              legend.width = 3.0,
              legend.format = list(digits = 6),
              legend.position = c("right", "top"),
              frame = TRUE) +
    tm_compass(type="8star", size = 2) +
    tm_scale_bar(width = 0.15) +
    tm_grid(lwd = 0.1, alpha = 0.5)
```

## SAM Method
```{r}
#| code-fold: true
#| code-summary: "Show the code"

exponent <- 2
method <- "SAM"

centroid.coords <- st_coordinates(st_centroid(grid))
points.coords <- st_coordinates(points)

dm <- exp(distance(centroid.coords, points.coords, type = "euclidean") / 1000 * exponent)

acc <- data.frame(ac(grid$demand,
                          points$capacity,
                          dm, 
                          d0 = 250,
                          power = 2, 
                          family = method))

colnames(acc) <- "acc"
hexagon <- bind_cols(grid, as_tibble(acc))
hexagon$acc[is.infinite(hexagon$acc)] <- NA

mapex <- st_bbox(grid)
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"

  tm_shape(hexagon,
           bbox = mapex) + 
    tm_fill(col = "acc",
            n = 5,
            style = "quantile",
            border.col = "black",
            border.lwd = 1,
            na.rm = TRUE) +
  tm_shape(points) +
    tm_symbols(size = 0.1) +
    tm_layout(main.title = paste("Accessibility to supermarkets: SAM method", sep=""),
              main.title.position = "center",
              main.title.size = 2,
              legend.outside = FALSE,
              legend.height = 0.45, 
              legend.width = 3.0,
              legend.format = list(digits = 6),
              legend.position = c("right", "top"),
              frame = TRUE) +
    tm_compass(type="8star", size = 2) +
    tm_scale_bar(width = 0.15) +
    tm_grid(lwd = 0.1, alpha = 0.5)
```
:::

## All-In One For Shiny

### Mapping to files
```{r}
#| code-fold: true
#| code-summary: "Show the code"

name2file <- new.env(hash=T, parent=emptyenv())

name2file[["markets and food centres"]] <- "markets_and_food_centres.rds"
name2file[["mrt"]] <- "mrt.rds"

name2file[["atm"]] <- "poi_atm.rds"
name2file[["bank"]] <- "poi_bank.rds"
name2file[["beauty salon"]] <- "poi_beauty_salon.rds"
name2file[["cafe"]] <- "poi_cafe.rds"
name2file[["clothing store"]] <- "poi_clothing_store.rds"
name2file[["convenience"]] <- "poi_convenience_store.rds"
name2file[["dentist"]] <- "poi_dentist.rds"
name2file[["doctor"]] <- "poi_doctor.rds"
name2file[["gym"]] <- "poi_gym.rds"
name2file[["hospital"]] <- "poi_hospital.rds"
name2file[["library"]] <- "poi_library.rds"
name2file[["lodging"]] <- "poi_lodging.rds"
name2file[["night club"]] <- "poi_night_club.rds"
name2file[["pow"]] <- "poi_place_of_worship.rds"
name2file[["restaurant"]] <- "poi_restaurant.rds"
name2file[["school"]] <- "poi_school.rds"
name2file[["tourist"]] <- "poi_tourist_attraction.rds"

name2file[["supermarkets"]] <- "supermarkets.rds"
name2file[["bus"]] <- "osm_sg.rds"
```

```{r}
#| code-fold: true
#| code-summary: "Show the code"

plot_acc <- function(method, quantiles, grid_size, point_type, exponent, subz) {
  grid <- read_rds(paste('data/rds/grid_', grid_size, '_hexagon', ifelse(subz, '_sz', '_pa'),'.rds', sep=""))
  
  points <- read_rds(paste('data/rds/', name2file[[point_type]], sep="")) %>%
    mutate(capacity = 500)
  
  centroid.coords <- st_coordinates(st_centroid(grid))
  points.coords <- st_coordinates(points)

  dm <- exp(distance(centroid.coords, points.coords, type = "euclidean") / 1000 * exponent)

  

  acc <- data.frame(ac(grid$demand,
                            points$capacity,
                            dm, 
                            d0 = 250,
                            power = 2, 
                            family = method))
  
  colnames(acc) <- "acc"
  hexagon <- bind_cols(grid, as_tibble(acc))
  hexagon$acc[is.infinite(hexagon$acc)] <- NA
  
  mapex <- st_bbox(grid)

  tm <- tm_shape(grid) + 
  tm_polygons() +
  tm_shape(hexagon,
           bbox = mapex) + 
    tm_fill(col = "acc",
            n = quantiles,
            style = "quantile",
            border.col = "black",
            border.lwd = 1,
            na.rm = TRUE) +
  tm_shape(points) +
    tm_symbols(size = 0.1) +
    tm_layout(main.title = paste("Accessibility to ", point_type, ": ", method," method", sep=""),
              main.title.position = "center",
              main.title.size = 1,
              legend.outside = FALSE,
              legend.height = 0.5, 
              legend.width = 0.5,
              legend.format = list(digits = 3),
              legend.position = c("right", "top"),
              frame = FALSE) +
    tm_compass(type="8star", size = 2) +
    tm_scale_bar(width = 0.20) +
    tm_grid(lwd = 0.1, alpha = 0.5)
  
  hexagon_acc <- st_join(hexagon, read_rds('data/rds/mpsz.rds') , join = st_intersects)
  
  region_bxp <- ggplot(data=hexagon_acc, 
       aes(y = acc, 
           x = REGION_N)) +
  geom_boxplot(outliers = FALSE) +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
  
  plot_grid(tmap_grob(tm), region_bxp, nrow = 2, rel_heights = c(2, 1))
}
```

https://stackoverflow.com/questions/66659389/combine-tmap-and-ggplot
```{r}
plot_acc("Hansen", 10, 500, "supermarkets", 2, TRUE)
```








































































